---
title: "pretty GPX Maps"
output: prettyGPXMaps
---

```{r}
source("setup_libraries.R")
source("code/utils.R")

# Stadia Maps API key (https://docs.stadiamaps.com/authentication/#api-keys)
stadiamaps_api_key <- Sys.getenv("STADIAMAPS_API_KEY")
```

## Settings

The map is created based on a single .gpx file. To visualize the combined track of multiple GPX files the merge function of [gpx studio](https://gpx.studio/app) can be used to obtain a single file. If the setting to keep the traces disconnected is used for merging, each sub-track of the merged file can be colored differently, by assigning a vector of colors to `track_colors` (see *GPXMapsMulticolor.RMD*).

Possible map styles: `stamen_toner_lite, stamen_toner, stamen_terrain, alidade_smooth, alidade_smooth_dark` (Note: only with the `'stamen'` style maps the track can be plotted below the map labels)

```{r}
# input file
gpx_file <- "gpx_files/Ligurien.gpx"

# set filenames for images that are created
map_filename <- "figures/ligurien_map_dark.png"
elevation_profile_filename <- "figures/ligurien_elevation.png"

# set map properties
map_style <- "stamen_toner"
start_end_markers <- FALSE
track_colors <- "#FC5200"

# set color for elevation profile plot
elevation_profile_color <- "#FC5200"
```

Load input GPX data

```{r}
gpx_track <- st_read(gpx_file, layer = "tracks", quiet = TRUE)
```

## Create Map

```{r}
map <- leaflet() |>
  addMapPane("base", zIndex = 400) |>
  addMapPane("track", zIndex = 500) |>
  addMapPane("labels", zIndex = 600) |>
  
  addStadiaMap(map_style, stadiamaps_api_key, pane = "base") |>
  addGpxTracks(gpx_track, track_colors, weight = 4, start_end_markers = start_end_markers, pane = "track") |>
  addStadiaMapLabels(map_style, stadiamaps_api_key, pane = "labels") |>
  setViewToGpx(gpx_track, zoom_offset = 0)  # set zoom_offset to get less/more detailed map

map
```

#### Save screenshot of map

Width and height might need to be adjusted based on the track shape.

Since the map tiles have a maximum resolution further increasing the zoom only improves the resolution the track is rendered at.

```{r}
mapshot(map, 
        file = map_filename,
        vwidth = 1000,
        vheight = 720,
        zoom = 8)  # zoom adjusts rendering resolution
```

## Create Elevation profile

```{r}
gpx_data_ele <- st_read(gpx_file, layer = "track_points", quiet = TRUE)
gpx_data_subset <- gpx_data_ele[round(seq(1, nrow(gpx_data_ele), length.out = 1000)), ]
base_level <- max(gpx_data_subset$ele)/25

elevation_plot <- ggplot(data = gpx_data_subset) +
  geom_area(aes(x = 1:nrow(gpx_data_subset), y = ele + base_level),
            fill = elevation_profile_color,
            alpha = 1) +
  theme_void()

elevation_plot
```

#### Save elevation profile

```{r}
ggsave(elevation_profile_filename,
       plot = elevation_plot,
       width = 20,
       height = 2,
       units = "cm",
       dpi = 900)
```
