---
title: "pretty GPX Maps"
output: prettyGPXMaps
---

```{r}
source("setup_libraries.R")
source("code/utils.R")
```

## Settings

Possible map styles: *`stamen_toner_lite, stamen_toner, stamen_terrain, alidade_smooth, alidade_smooth_dark`*

Only with `stamen` style maps the track can be plotted below the map labels.

```{r}
# input file
gpx_file <- "gpx_files/Ligurien.gpx"

# set filenames for images that are created
map_filename <- "figures/ligurien_map_dark.png"
elevation_profile_filename <- "figures/ligurien_elevation.png"

# set map properties
map_style <- "stamen_toner"
start_end_markers <- FALSE
track_color <- "#FC5200"

# set color for elevation profile plot
elevation_profile_color <- "#FC5200"

# Stadia Maps API key (https://docs.stadiamaps.com/authentication/#api-keys)
stadiamaps_api_key <- Sys.getenv("STADIAMAPS_API_KEY")
```

Load input GPX data

```{r}
gpx_track <- st_read(gpx_file, layer = "tracks", quiet = TRUE)
```

## Create Map

```{r}
map <- leaflet() |>
  addMapPane("base", zIndex = 400) |>
  addMapPane("track", zIndex = 500) |>
  addMapPane("labels", zIndex = 600) |>
  
  addStadiaMap(map_style, stadiamaps_api_key, pane = "base") |>
  addPolylines(data = gpx_track, color = track_color, weight = 4, opacity = 1, smoothFactor = 1, options = pathOptions(pane = "track")) |>
  addStadiaMapLabels(map_style, stadiamaps_api_key, pane = "labels") |>
  setViewToGpx(gpx_track)
  
if (start_end_markers) {
  map <- addStartEndMarkers(map, gpx_track, track_color)
}

map
```

#### Save screenshot of map

Width and height might need to be adjusted based on the track shape.

Since the map tiles have a maximum resolution further increasing the zoom only improves the resolution the track is rendered at.

```{r}
mapshot(map, 
        file = map_filename,
        vwidth = 1000,
        vheight = 720,
        zoom = 8)  # zoom adjusts rendering resolution
```

## Elevation profile

```{r}
gpx_data_ele <- st_read(gpx_file, layer = "track_points", quiet = TRUE)
gpx_data_subset <- gpx_data_ele[round(seq(1, nrow(gpx_data_ele), length.out = 1000)), ]
base_level <- max(gpx_data_subset$ele)/25

elevation_plot <- ggplot(data = gpx_data_subset) +
  geom_area(aes(x = 1:nrow(gpx_data_subset), y = ele + base_level),
            fill = elevation_profile_color,
            alpha = 1) +
  theme_void()

elevation_plot
```

Save elevation profile

```{r}
ggsave(elevation_profile_filename,
       plot = elevation_plot,
       width = 20,
       height = 2,
       units = "cm",
       dpi = 900)
```
