---
title: "pretty GPX Maps"
output: prettyGPXMaps
---

```{r}
source("setup_libraries.R")
source("code/utils.R")
```

## Settings

The map is created based on a single .gpx file. To visualize the combined track of multiple GPX files the merge function of [gpx studio](https://gpx.studio/app) can be used to obtain a single file. If the setting to keep the traces disconnected is used for merging, each sub-track of the merged file can be colored differently, by assigning a vector of colors to `track_colors`.

Possible map styles: `stamen_toner_lite, stamen_toner, stamen_terrain, alidade_smooth, alidade_smooth_dark` (Note: only with the `'stamen'` style maps the track can be plotted below the map labels)

```{r}
# input file
gpx_file <- "gpx_files/Karlsfelder_Triathlon.gpx"

# set filename for map that is created
map_filename <- "figures/triathlon_karlsfeld.png"

# set map properties
map_style <- "stamen_toner"
start_end_markers <- FALSE
# colors for swim, transition 1, bike, transition 2, run
track_colors <- c("#007AFF", "#000000", "#34C759", "#000000", "#FF2D55")  

# Stadia Maps API key (https://docs.stadiamaps.com/authentication/#api-keys)
stadiamaps_api_key <- Sys.getenv("STADIAMAPS_API_KEY")
```

Load input GPX data

```{r}
gpx_track <- st_read(gpx_file, layer = "tracks", quiet = TRUE)
```

## Create Map

```{r}
map <- leaflet() |>
  addMapPane("base", zIndex = 400) |>
  addMapPane("track", zIndex = 500) |>
  addMapPane("labels", zIndex = 600) |>
  
  addStadiaMap(map_style, stadiamaps_api_key, pane = "base") |>
  addGpxTracks(gpx_track, track_colors, weight = 4, start_end_markers = start_end_markers, pane = "track") |>
  addStadiaMapLabels(map_style, stadiamaps_api_key, pane = "labels") |>
  setViewToGpx(gpx_track, zoom_offset = 0)  # set zoom_offset to get less/more detailed map

map
```

#### Save screenshot of map

Width and height might need to be adjusted based on the track shape. Since the map tiles have a maximum resolution further increasing the zoom only improves the resolution the track is rendered at.

```{r}
mapshot(map, 
        file = map_filename,
        vwidth = 1000,
        vheight = 500,
        zoom = 8)  # zoom adjusts rendering resolution
```
